% \VignettePackage{switchr}
% \VignetteIndexEntry{switchr basics}
% \VignetteEngine{knitr::rmarkdown}

# The switchr package

## Managing and populating package libraries


Different versions of R packages are appropriate for different applications,
including package development, solo analysis, and
collaboration. *switchr* provides a framework for creating, populating,
managing, and switching between multiple package libraries from within
an R session.

```{r echo=FALSE, results="hide"}
library(switchr)
switchrBaseDir(file.path(tempdir(), ".switchr"))
switchrDontUnload(c("knitr", "evaluate", "digest", "formatR", "highr", 
                    "markdown", "stringr"))
options(width=40)

```


We create and use switchr package libraries by referring to them by
name using the /switchTo()/ function as seen below.

```{r}
switchTo("vign1")
```

The first time this code is run a "vign1" library is created and activated. By
activate we mean that it replaces the current User- and (by default) Site- 
library locations which R uses to find packages during the loading process.

To stop using "vign1" we simply call /switchBack()/:

```{r}
switchBack()
```


## Representing package cohorts

Empty package libraries are, of course, of limited use. The other main 
functionality *switchr* offers is installation of packages from a wide
variety of sources (git, svn, package repositories, and the CRAN Archive
among them), both individually and as cohorts.

We do this via a generalized form of a package repository called a 
/package manifest/, which includes a listing of packages and the information
switchr needs to install them (primarily location of their source code).

We can create a package manifest from scratch via the PkgManifest, though there
convenience functions for common use-cases such as Github packages

```{r}
man = PkgManifest(name = "fastdigest", 
    url = "https://github.com/gmbecker/fastdigest")
man
```

The `GithubManifest` convenience function builds manifests from GitHub username/repository
pairs for package sources residing in the root directory of the master branch for the 
specified repository.
```{r}
man2 = GithubManifest("gmbecker/fastdigest", 
     "gmbecker/RCacheSuite")
man2
```

## Installing packages and package cohorts

*switchr* provides the /install_packages/ function for installing specific packages
or sets of packages from a variety of sources, including traditional repositories
and package manifests. This function supports package dependencies, including 
non-repository to non-repository dependecies (e.g. between two or more packages
on GitHub) on systems able to build packages from source.



NOT RUN
```{r eval=FALSE}
install_packages("RCacheSuite", man2)
```

