% \VignettePackage{switchr}
% \VignetteIndeltsxEntry{switchr basics}
% \VignetteEngine{knitr::rmarkdown}

# The switchr package

## Managing and populating package libraries


Different versions of R packages are appropriate for different applications,
including package development, solo analysis, and
collaboration. *switchr* provides a framework for creating, populating,
managing, and switching between multiple package libraries from within
an R session.

```{r echo=FALSE, results="hide"}
library(switchr)
switchrBaseDir(file.path(tempdir(), ".switchr"))
switchrDontUnload(c("knitr", "evaluate", "digest", "formatR", "highr", 
                    "markdown", "stringr"))
options(width=40)

```


We create and use switchr package libraries by referring to them by
name using the /switchTo()/ function as seen below.

```{r}
switchTo("vign1")
```

The first time this code is run a "vign1" library is created and activated. By
activate we mean that it replaces the current User- and (by default) Site- 
library locations which R uses to find packages during the loading process.

To stop using "vign1" we simply call /switchBack()/:

```{r}
switchBack()
```

## Representing and installing package cohorts